Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2019-02-14 WholePlotTrajectories.Rmd
files

The large areas covered with tropical forests worldwide hold crucial environmental, economic and social values. 
They provide wood and multiple non-timber forest products, shelter a diversified fauna, and ensure cultural and human well-being.
They regulate as well the local and regional climates, and the carbon, water and nutrient cycles.
However, the growing demand in forests products together with current global changes increase the pressure on remaining natural forests [@Morales-Hidalgo2015].
These threats affect the natural disturbance regime that shape and maintain the structure, composition and functioning of tree communities [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
It is now urgent to understand the processes underlying community response to disturbance, and thus to be able to anticipate the fate of tropical forests in the current context.
In tropical forests, disturbance maintains a forest environment and initiates a suit of changes in the structure and functioning of communities consituting a secondary succession.
Disturbance change both community abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017]. 
For now, post-disturbance trajectories has been largely studied through forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
Models based on observed trajectories then assessed community time to recovery and identified some of the determinants of post-disturbance trajectories, like forest structure, forest composition, environmental parameters and disturbance characteristics [@Herault2018]. 
Regarding community diversity and composition, however, post-disturbance trajectories have not been as thoroughly understood [@Guitet2018;@Molino2001].
Given the variety of species response to disturbance and the huge diversity of tropical forests, biodiversity trajectories following disturbance might be very diverse [@Lindenmayer2012;@Garcia_florez2017].
We assumed that the rate, endpoint and duration of the trajectories depend on the intensity and frequency of disturbance: we then aimed to highlight the link between biodiversity and disturbance and thus anticipate community post-disturbance trajectories [@Chazdon2003a;@Willig2018].

An early conceptual basis of the linkage between biodiversity and disturbance is the Intermediate Disturbance Hypothesis (IDH).
The IDH assumes a relationship between community diversity and the intensity and frequency of disturbance events [@Connell1978].
The theory and states that community diversity peaks at intermediate level of disturbance.
The fluctuations of environmental conditions after disturbance are assumed to prevent competitive exclusion in fostering both competitively superior species and fast colonizers [@Shea2004;@Pulsford2016].
In highly diverse tropical forests, however, observations of the IDH are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
The scope of the IDH should then be clarified regarding the different facets of community richness, abundance distribution and functioning.
The underlying processes of the IDH are besides part of the post-disturbance succession and the question is on the time after disturbance validating the hypothesis.
In this study we aimed as well to clarify the scope of the IDH and discuss its implication for the fate of tropical forests [@Sheil2003;@Shea2004].

>>>>

The large areas covered with tropical forests worldwide hold crucial environmental, economic and social values. 
They provide wood and multiple non-timber forest products, shelter a diversified fauna, and ensure cultural and human well-being.
They regulate as well the local and regional climates, and the carbon, water and nutrient cycles.
However, the growing demand in forests products together with current global changes increase the pressure on remaining natural forests [@Morales-Hidalgo2015].
These threats affect the natural disturbance regime that shape and maintain the structure, composition and functioning of tree communities [@Schnitzer2001;@Anderson-Teixeira2013;@Sist2015].
It is urgent to understand community response to disturbance and the processes underlying to anticipate the fate of tropical forests in the current context.
Disturbance change both community abiotic and biotic environments, through modifications in the fluxes of light, heat and water [@Goulamoussene2017]. 
In tropical forests, the forest environment is maintained after disturbance and there is a suit of changes in the structure and functioning of communities that consitutes a secondary succession.
For now, this succession has been largely studied through post-disturbance trajectories of forest structural parameters such as aboveground biomass, tree height or stem density [@Piponiot2016;@Rutishauser2016]. 
Models based on observed trajectories then assessed community time to recovery and identified some of the determinants of post-disturbance trajectories, like forest structure, forest composition, environmental parameters and disturbance characteristics [@Herault2018]. 
Regarding community diversity and composition, however, post-disturbance trajectories have not been as thoroughly understood [@Guitet2018;@Molino2001].
Given the variety of species response to disturbance and the huge diversity of tropical forests, post-disturbance biodiversity trajectories might be very diverse [@Lindenmayer2012;@Garcia_florez2017].
It is essential to anticipate these trajectories, and hence to highlight the link between biodiversity and disturbance, specifically how disturbance intensity and frequency determine the rate, endpoint and duration of the trajectories [@Chazdon2003a;@Willig2018].

An early conceptual basis of the linkage between biodiversity and disturbance is the Intermediate Disturbance Hypothesis (IDH).
The IDH assumes a relationship between community diversity and the intensity and frequency of disturbance events [@Connell1978].
The theory and states that community diversity peaks at intermediate level of disturbance.
The fluctuations of environmental conditions after disturbance are assumed to prevent competitive exclusion in fostering both competitively superior species and fast colonizers [@Shea2004;@Pulsford2016].
In highly diverse tropical forests, however, observations of the IDH are often blurry compared to theoretical expectations [@Hugues2007;@Sheil2003;@Norden2017], and the validation of the theory remains controversial [@Hubbell2001;@Fox2013;@Sheil2013].
Given the huge diversity of tropical forests and the variety of ecological processes that may underlie the IDH The scope of the IDH need be clarified regarding the different facets of community richness, abundance distribution and functioning.
The underlying processes of the IDH are besides part of the post-disturbance succession and the question is on the time after disturbance validating the hypothesis.
In this study we aimed as well to clarify the scope of the IDH and discuss its implication for the fate of tropical forests [@Sheil2003;@Shea2004].
